Source-dependent likelihood-ratio ranking statistic for compact binary coalescence candidates

### 2016-09-30 article_rankstatsrc.tex
moved commented section to end of paper
explained colour plots more

In Section \ref{sec:rankingstat}, we begin by describing the ranking statistic that is currently used in Advanced LIGO's low-latency searches for compact binary coalescences. The derivation of a ranking statistic that envelops astrophysical source model information is outlined in Section \ref{sec:derivation}, where the various terms are explored in further detail in Sections \ref{sec:P_j} and \ref{sec:P_k}. In Section \ref{sec:P_jk}, we present a discussion of the source-dependent ranking statistic, and conclude in Section \ref{sec:conclusion}.

@@

\begin{figure}
\resizebox{\linewidth}{!}
{\includegraphics{logPj_OzelrhoNone.png}}
\resizebox{\linewidth}{!}
{\includegraphics{logPj_BBHrhoNone.png}}
\caption{The logarithmic probability of the signal population, i.e. the probability $P_j$ that a signal $t_j$ exists. Only $m_1$ and $m_2$ of the template bank are plotted. We use two different signal populations: BNS (top) and BBH (bottom). The black solid lines show the mass distribution of the different population models. Note that log$P_j=-\infty$ is not plotted.}
\label{fig:p_j}
\end{figure}

@@

\FIXME{Replace Kiziltan plots with BBH}

@@

The probabilities currently computed are a function of $\rho$: $P(\text{a signal is recovered by template }j|\rho)$. The values of $\rho$ are chosen via Chebyshev nodes, as these are shown to reduce Runge's phenomenon when fitting \cite{numericalmethodsMATLAB}. The Chebyshev nodes in the interval (-1,1) are determined by the following:
	\begin{equation}
	x_k = \cos\Big(\frac{2k-1}{2n}\pi\Big), \quad k=1,...,n
	\end{equation}
where $n$ is a natural number.

@@

We assume that $P(\rho)\propto\rho^{-4}$. We cannot normalize this because there would be an infinite density of 0 SNR sources at infinite distance. As a result, we pick the lowest SNR to be $\rho_{\text{low}}=4$ as this is the trigger threshold for a gravitational-wave search, and we integrate to $\rho_{\text{high}}=\infty$.

\begin{figure}
\includegraphics[width=0.5\textwidth]{logP_Ozelrho0_0.png}
\includegraphics[width=0.5\textwidth]{logP_Ozelrho16_3802850078.png}
\includegraphics[width=0.5\textwidth]{logP_Ozelrho51_2132034356.png}
\caption{$P(\vec{t}_k|\text{signal},\rho)$ for different values of $\rho$ for the BNS source population. The solid black lines denote the mass distribution of the respective source population model.}
\end{figure}
\begin{figure}
\includegraphics[width=0.5\textwidth]{logP_Kiziltanrho0_0.png}
\includegraphics[width=0.5\textwidth]{logP_Kiziltanrho16_3802850078.png}
\includegraphics[width=0.5\textwidth]{logP_Kiziltanrho51_2132034356.png}
\caption{$P(\vec{t}_k|\text{signal},\rho)$ for different values of $\rho$ for the BBH source population. The solid black lines denote the mass distribution of the respective source population model.}
\end{figure}

\begin{figure}
\includegraphics[width=0.5\textwidth]{lagrangian_fit_Ozel.pdf}
\includegraphics[width=0.5\textwidth]{lagrangian_fit_Kiziltan.pdf}
\caption{Lagrangian fits of $P_{jk}$ vs $\rho$ for a few templates (top: BNS, bottom: BBH). Dots are the data points, and solid lines are the functions found fitting Lagrangian polynomials to the data \FIXME{(change to cubic splines)}. The $\rho$ values are found using the Chebychev nodes, with the addition of $\rho=0$.}
\label{fig:Pvsrho}
\end{figure}

To better understand which templates give asymptotically approach a constant value with increasing $\rho$, in Figure~\ref{fig:P_tkj}, we show the templates mapped onto mass axes, with the colours denoting the magnitude of the calculated probability. As $\rho$ increases, we see that the probabilities of most templates exponentially decrease, except for a collection of templates that are clustered around the source population.

@@

%%%%%%%%%%%%%%%%%%%%%%
\bibliography{references}
\end{document}

>>>>

In Section \ref{sec:rankingstat}, we begin by describing the ranking statistic that is currently used in Advanced LIGO's low-latency searches for compact binary coalescences. The derivation of a ranking statistic that envelops astrophysical source model information is outlined in Section \ref{sec:derivation}, where the various terms are explored in further detail in Sections \ref{sec:P_j} and \ref{sec:P_k}. In Section \ref{sec:P_jk}, we present a discussion of the source-dependent ranking statistic and test it using a BNS and BBH source population model. We conclude in Section \ref{sec:conclusion}.

@@

\begin{figure}
\resizebox{\linewidth}{!}
{\includegraphics{logPj_OzelrhoNone.png}}
\resizebox{\linewidth}{!}
{\includegraphics{logPj_BBHrhoNone.png}}
\caption{The logarithmic probability of the signal population, i.e. the probability $P_j$ that a signal $t_j$ exists. Only $m_1$ and $m_2$ of the template bank are plotted. We use two different signal populations: BNS (top) and BBH (bottom). The black solid lines show the mass distribution of the different population models. Note log$P_j=-\infty$ is not plotted.}
\label{fig:p_j}
\end{figure}

@@

@@

The probabilities currently computed are a function of $\rho$: $P(\text{a signal is recovered by template }j|\rho)$. The values of $\rho$ are chosen via Chebyshev nodes, as these are shown to reduce Runge's phenomenon when fitting \cite{numericalmethodsMATLAB}. The Chebyshev nodes in the interval (-1,1) are determined by the following:
	\begin{equation}
	x_k = \cos\Big(\frac{2k-1}{2n}\pi\Big), \quad k=1,...,n
	\end{equation}
where $n$ is a natural number. We scale and shift these nodes such that $0\le\rho\le60$.

@@

We assume that $P(\rho)\propto\rho^{-4}$. We cannot normalize this because there would be an infinite density of 0 SNR sources at infinite distance. Instead, we normalize $\rho^{-4}$, from $\rho_{\text{low}}=4$, the trigger threshold for a gravitational-wave search, to $\rho_{\text{high}}=\infty$.

\begin{figure}
\includegraphics[width=0.5\textwidth]{logP_Ozelrho0_0.png}
\includegraphics[width=0.5\textwidth]{logP_Ozelrho16_3802850078.png}
\includegraphics[width=0.5\textwidth]{logP_Ozelrho51_2132034356.png}
\caption{log($P(\vec{t}_k|\text{signal},\rho)$) for different values of $\rho$ for the BNS source population. The solid black lines denote the mass distribution of the respective source population model.}
\label{fig:P_tkj_BNS}
\end{figure}
\begin{figure}
\includegraphics[width=0.5\textwidth]{logP_Kiziltanrho0_0.png}
\includegraphics[width=0.5\textwidth]{logP_Kiziltanrho16_3802850078.png}
\includegraphics[width=0.5\textwidth]{logP_Kiziltanrho51_2132034356.png}
\caption{log($P(\vec{t}_k|\text{signal},\rho)$) for different values of $\rho$ for the \FIXME{Kiziltan BNS - change to BBH} source population. The solid black lines denote the mass distribution of the respective source population model.}
\label{fig:P_tkj_BBH}
\end{figure}

\begin{figure}
\includegraphics[width=0.5\textwidth]{lagrangian_fit_Ozel.pdf}
\includegraphics[width=0.5\textwidth]{lagrangian_fit_Kiziltan.pdf}
\caption{Lagrangian fits of $P_{jk}$ vs $\rho$ for a few templates (top: BNS, bottom: \FIXME{Kiziltan BNS - change to BBH}). Dots are the data points, and solid lines are the functions found using a \FIXME{Lagrangian fit - change to cubic interpolation with increased $\rho$ sampling}. The $\rho$ values are found using the Chebychev nodes, with the addition of $\rho=0$.}
\label{fig:Pvsrho}
\end{figure}

To better understand which templates asymptotically approach a constant value with increasing $\rho$, in Figures~\ref{fig:P_tkj_BNS} and \ref{fig:P_tkj_BBH}, we show the templates mapped onto mass axes, with the colours denoting the magnitude of the calculated logarithmic probability. As $\rho$ increases, we see that the probabilities of most templates exponentially decrease, except for a collection of templates that are clustered around the source population. At high $\rho$, the probability distribution matches that of the source population distribution. This is expected, as for extremely strong signals, it means that the template that recovers the signal must be equal to the signal template itself. However, at low $\rho$, we see that probabilities are high for templates along the line of constant chirp mass. These templates have high overlaps with each other and are also likely to recover signals.

@@

\bibliography{references}
\end{document}
